```{r include_packages_2, include = FALSE}
# This chunk ensures that the thesisdown package is
# installed and loaded. This thesisdown package includes
# the template files for the thesis and also two functions
# used for labeling and referencing
if(!require(devtools))
  install.packages("devtools", repos = "http://cran.rstudio.com")
if(!require(dplyr))
    install.packages("dplyr", repos = "http://cran.rstudio.com")
if(!require(ggplot2))
    install.packages("ggplot2", repos = "http://cran.rstudio.com")
if(!require(ggplot2))
    install.packages("bookdown", repos = "http://cran.rstudio.com")
if(!require(thesisdown)){
  library(devtools)
  devtools::install_github("ismayc/thesisdown")
  }
library(thesisdown)
flights <- read.csv("data/flights.csv")
```

# Data Analysis
## Change in Largest Holdings Over Time
The top 5 largest holdings in the US Equal Weight Index by average weight are Apple (AAPL), Exxon Mobil (XOM), Microsoft (MSFT), General Electric (GE), and Johnson & Johnson (JNJ). Each stock's average weight is shown below in Table 4.1. 

```{r, echo = FALSE}
load("~/thesis_final/data/usa.Rda")

usa_topweight <- aggregate(weight ~ ticker, data=usa, FUN=sum)
usa_topweight <- arrange(usa_topweight, desc(weight))
usa_topweight$weight <- usa_topweight$weight/63
usa_topweight$rank <- seq.int(nrow(usa_topweight))
```

```{r fig.cap = "Top 5 Holdings in US Equal Weight Index over Time by Average Weight", echo=FALSE}

usa_topweight <- select(usa_topweight, rank, ticker, weight)
usa_topweight <- usa_topweight[1:5,]

kable(usa_topweight, 
      col.names = c("Rank", "Ticker", "Average Weight (%)"),
      caption = "Top 5 Holdings in the US Equal Weight Index over Time by Average Weight",
      longtable = TRUE,
      booktabs = TRUE)
```

The top 5 largest holdings in the Minimum Volatility Index by average weight are Verizon (VZ), AT&T (T), Automatic Data Processing (ADP), Johnson & Johnson (JNJ), and McDonald's (MCD). Each stock's average weight is shown below in Table 4.2. 

```{r, echo = FALSE}
load("~/thesis_final/data/minvol.Rda")

minvol_topweight <- aggregate(weight ~ ticker, data=minvol, FUN=sum)
minvol_topweight <- arrange(minvol_topweight, desc(weight))
minvol_topweight$weight <- minvol_topweight$weight/63
minvol_topweight$rank <- seq.int(nrow(minvol_topweight))
```

```{r fig.cap = "Top 5 Holdings in Minimum Volatility Index over Time by Average Weight", echo=FALSE}

minvol_topweight <- select(minvol_topweight, rank, ticker, weight)
minvol_topweight <- minvol_topweight[1:5,]

kable(minvol_topweight, 
      col.names = c("Rank", "Ticker", "Average Weight (%)"),
      caption = "Top 5 Holdings in the Minimum Volatility Index over Time by Average Weight",
      longtable = TRUE,
      booktabs = TRUE)
```

The 5 largest holdings in the US Equal Weight index are AAPLE, XOM, MSFT, GE, and JNJ. As shown in Figure \ref{fig:plot5}, the weights of the 5 companies start off very high, each comprising a few percent of the overall index, then suddenly all drop significantly to a fraction of a percent after 2015-08-31. After verifying this with iShares and MSCI, it was discovered that this due to change in the weighting mechanism.\newline
```{r, echo = FALSE}
load("~/thesis_final/data/usa.Rda")
load("~/thesis_final/data/minvol.Rda")

usa_sub1 <- filter(usa, ticker == "AAPL" | ticker == "XOM" | ticker == "MSFT" | ticker == "GE" | ticker == "JNJ")
usa_sub1 <- select(usa_sub1, date, ticker, weight)

library(ggplot2)

holdings1<- ggplot() + 
	geom_line(data = filter(usa_sub1, ticker == "AAPL"), aes(x = date, y = weight, color = "AAPL")) +
	geom_line(data = filter(usa_sub1, ticker == "XOM"), aes(x = date, y = weight, color = "XOM"))  +
	geom_line(data = filter(usa_sub1, ticker == "MSFT"), aes(x = date, y = weight, color = "MSFT"))  +
	geom_line(data = filter(usa_sub1, ticker == "GE"), aes(x = date, y = weight, color = "GE"))  +
	geom_line(data = filter(usa_sub1, ticker == "JNJ"), aes(x = date, y = weight, color = "JNJ"))  +
	xlab('Date') + ylab('Weight (%)') + ggtitle('Average Weight Change of Top 5 Holdings in the US Equal Weight Index')
```

``` {r, fig.cap = "Change in weight for the top 5 holdings in the US Equal Weight index over time\\label{fig:plot5}", fig.height=2.8, fig.width=6.5, echo=FALSE}
plot(holdings1)
```

\newline
The 5 largest holdings of the Minimum Volatility index are VZ, T, ADP, JNJ, and MCD. As shown in Figure \ref{fig:plot6}, since the index's inception these stocks have generally oscillated between a weight of 1.3% and 1.7% of the overall portfolio, with the exception of Verizon which reached around 2% in 2014.
```{r, echo = FALSE}
load("~/thesis_final/data/minvol.Rda")
minvol_sub1 <- filter(minvol, ticker == "VZ" | ticker == "T" | ticker == "ADP" | ticker == "JNJ" | ticker == "MCD")
minvol_sub1 <- select(minvol_sub1, date, ticker, weight)

library(ggplot2)

holdings2<- ggplot() + 
	geom_line(data = filter(minvol_sub1, ticker == "VZ"), aes(x = date, y = weight, color = "VZ")) +
	geom_line(data = filter(minvol_sub1, ticker == "T"), aes(x = date, y = weight, color = "T"))  +
	geom_line(data = filter(minvol_sub1, ticker == "ADP"), aes(x = date, y = weight, color = "ADP"))  +
	geom_line(data = filter(minvol_sub1, ticker == "JNJ"), aes(x = date, y = weight, color = "JNJ"))  +
	geom_line(data = filter(minvol_sub1, ticker == "MCD"), aes(x = date, y = weight, color = "MCD"))  +
	xlab('Date') + ylab('Weight (%)') + ggtitle('Average Weight Change of Top 5 Holdings in the Minimum Volatility Index')
```

``` {r fig.cap = "Change in weight for the top 5 holdings in the Minimum Volatility index over time\\label{fig:plot6}", fig.height=2.8, fig.width=6.5, echo=FALSE}
plot(holdings2)
```

\clearpage

## Change in Sector Weights Over Time
Sector weights were calculated over time for both the US Equal Weight and the Minimum Volatility Index by dividing the number of holdings for each industry by the total number of holdings in the portfolio. This was done to get a sense of what industries may be inherently more "low-risk" or "high-risk." Moreover, this could help verify the accuracy of the data, as each sector weighting in the Minimum Volatility index should be within 5% of each sector weighting in the US Equal Weight Index, as specified by the Barra Open Optimizer [@msci2013].

Shown below in Figure \ref{fig:sector1} are the sector weights over time for energy stocks in both indices. The weighting of energy stocks in the US Equal Weight Index is consistently greater than the weighting of energy stocks in the Minimum Volatility Index. This could imply that by nature, energy stocks are more volatile than stocks in other industries. 

```{r, echo=FALSE}
load("~/thesis_final/data/usa_percent.Rda")
load("~/thesis_final/data/minvol_percent.Rda")
library(ggplot2)
```

``` {r fig.cap = "Weight of energy stocks by index over time\\label{fig:sector1}", fig.height=4.5, fig.width=6.5, echo=FALSE}
## Energy
Eng1 <- usa_percent[which(usa_percent$sector_name=="Energy"), ]
Eng2 <- minvol_percent[which(minvol_percent$sector_name=="Energy"), ]
ggplot(Eng1, aes(date, percent, colour = "US Equal Weight")) + geom_line() +  
ggtitle("Energy Sector Weight Over Time") + xlab("Time") + ylab("Sector Weight") +
geom_line(data = Eng2, aes(x=date, y=percent, colour="Minimum Volatility"),show.legend = TRUE)
```

\clearpage
Shown below in Figure \ref{fig:sector2} are the sector weights over time for financial stocks in both indices. The weighting of financial stocks in the US Equal Weight Index is fairly consistent, until a significant drop towards the latter half of 2016. This timing coincides with the Federal Reserve's decision to hike rates for just the second time in a decade, amid positive economic signs [@cox2016]. In the Minimum Volatility Index, the weighting of financials fluctuate quite a bit, and experience a similar drop in weight at that same point in time.

``` {r fig.cap = "Weight of financial stocks by index over time\\label{fig:sector2}", fig.height=2.8, fig.width=6.5, echo=FALSE}
## Financials
Fin1 <- usa_percent[which(usa_percent$sector_name=="Financials"), ]
Fin2 <- minvol_percent[which(minvol_percent$sector_name=="Financials"), ]
ggplot(Fin1, aes(date, percent, colour = "US Equal Weight")) + geom_line() + 
ggtitle("Financial Sector Weight Over Time")+xlab("Time")+ylab("Sector Weight") + 
geom_line(data = Fin2, aes(x=date, y=percent, colour="Minimum Volatility"),show.legend = TRUE)
```

\newline
Shown below in Figure \ref{fig:sector3} are the sector weights over time for consumer staple stocks in both indices. The weighting of consumer staples in the US Equal Weight Index is consistently lower than the weighting of consumer staples in the Minimum Volatility Index. This could imply that consumer staple stocks are less volatile than stocks in other industries. 

``` {r fig.cap = "Weight of consumer staple stocks by index over time\\label{fig:sector3}", fig.height=2.8, fig.width=6.5, echo=FALSE}
## Consumer Staples
ConStap1 <- usa_percent[which(usa_percent$sector_name=="Consumer Staples"), ]
ConStap2 <- minvol_percent[which(minvol_percent$sector_name=="Consumer Staples"), ]
ggplot(ConStap1, aes(date, percent, colour = "US Equal Weight")) + geom_line() + 
ggtitle("Consumer Staples Sector Weight Over Time") + xlab("Time") + 
ylab("Sector Weight") + geom_line(data = ConStap2, aes(x=date, y=percent, 
colour="Minimum Volatility"),show.legend = TRUE)
```

\clearpage
Shown below in Figure \ref{fig:sector4} are the sector weights over time for consumer discretionary stocks in both indices. The weight of consumer discretionary stocks is consistently greater in the US Equal Weight Index than in the Minimum Volatility Index. This could imply that consumer staple stocks are more volatile than stocks in other industries. 
``` {r, cache=TRUE, echo=FALSE}
## Consumer Discretionary
ConDis1 <- usa_percent[which(usa_percent$sector_name=="Consumer Discretionary"), ]
ConDis2 <- minvol_percent[which(minvol_percent$sector_name=="Consumer Discretionary"), ]
cons1<- ggplot(ConDis1, aes(date, percent, colour = "US Equal Weight")) + geom_line() + 
ggtitle("Consumer Discretionary Sector Weight Over Time") + xlab("Time") + 
ylab("Sector Weight") + geom_line(data = ConDis2, aes(x=date, y=percent, 
colour="Minimum Volatility"),show.legend = TRUE)
```


``` {r fig.cap = "Weight of consumer discretionary stocks by index over time\\label{fig:sector4}", fig.height=2.85, fig.width=6.5, echo=FALSE}
plot(cons1)
```

\newline
Shown below in Figure \ref{fig:sector5} are the sector weights over time for healthcare stocks in both indices. The weight of healthcare stocks is consistently higher over time in the Minimum Volatility Index than in the US Equal Weight Index. This could imply that consumer staple stocks are less volatile than stocks in other industries. 

``` {r, cache=TRUE, echo=FALSE}
## Healthcare
Health1 <- usa_percent[which(usa_percent$sector_name=="Health Care"), ]
Health2 <- minvol_percent[which(minvol_percent$sector_name=="Health Care"), ]
healthcare1<-ggplot(Health1, aes(date, percent, colour = "US Equal Weight")) + geom_line() + 
ggtitle("Healthcare Sector Weight Over Time") + xlab("Time") + 
ylab("Sector Weight") + geom_line(data = Health2, aes(x=date, y=percent, 
colour="Minimum Volatility"),show.legend = TRUE)
```

``` {r fig.cap = "Weight of healthcare stocks by index over time\\label{fig:sector5}", fig.height=2.85, fig.width=6.5, echo=FALSE}
plot(healthcare1)
```


\clearpage
Shown below in Figure \ref{fig:sector6} are the sector weights over time for industrial stocks in both indices. The weight of industrials is consistently higher over time in the US Equal Weight Index than in the Minimum Volatility Index. This could imply that industrial stocks are more volatile than stocks in other industries. 
``` {r, cache=TRUE, echo=FALSE}
Ind1 <- usa_percent[which(usa_percent$sector_name=="Industrials"), ]
Ind2 <- minvol_percent[which(minvol_percent$sector_name=="Industrials"), ]
industrials1<-ggplot(Ind1, aes(date, percent, colour = "US Equal Weight")) + geom_line() + 
ggtitle("Industrials Sector Weight Over Time") + xlab("Time") + 
ylab("Sector Weight") + geom_line(data = Ind2, aes(x=date, y=percent, 
colour="Minimum Volatility"),show.legend = TRUE)
```

``` {r fig.cap = "Weight of industrials stocks by index over time\\label{fig:sector6}", fig.height=2.85, fig.width=6.5, echo=FALSE}
plot(industrials1)
```

\newline
Shown below in Figure \ref{fig:sector7} are the sector weights over time for information technology stocks in both indices. The weight of information technology stocks between 2012 and 2013 was significantly greater in the Minimum Volatility Index than in the US Equal Weight Index. However, since 2014, the weights of IT in both indices have converged, implying the industry may have become more volatile over the last few years.

``` {r, cache=TRUE, echo=FALSE}
## Information Technology
IT1 <- usa_percent[which(usa_percent$sector_name=="Information Technology"), ]
IT2 <- minvol_percent[which(minvol_percent$sector_name=="Information Technology"), ]
infotech1<- ggplot(IT1, aes(date, percent, colour = "US Equal Weight")) + geom_line() + 
ggtitle("Information Technology Sector Weight Over Time") + 
xlab("Time") + ylab("Sector Weight") + geom_line(data = IT2, aes(x=date, y=percent,
colour="Minimum Volatility"),show.legend = TRUE)
```

``` {r fig.cap = "Weight of information technology stocks by index over time\\label{fig:sector7}", fig.height=2.85, fig.width=6.5, echo=FALSE}
plot(infotech1)
```

\clearpage
Shown below in Figure \ref{fig:sector8} are the sector weights over time for materials stocks in both indices. The weight of materials is consistently higher over time in the US Equal Weight Index than in the Minimum Volatility Index. This could imply that materials companies may be more volatile than companies in other sectors.
``` {r, cache=TRUE, echo=FALSE}
## Materials
Mat1 <- usa_percent[which(usa_percent$sector_name=="Materials"), ]
Mat2 <- minvol_percent[which(minvol_percent$sector_name=="Materials"), ]
materials1<- ggplot(Mat1, aes(date, percent, colour = "US Equal Weight")) + geom_line() + 
ggtitle("Materials Sector Weight Over Time") + xlab("Time") + 
ylab("Sector Weight") + geom_line(data = Mat2, aes(x=date, y=percent, 
colour="Minimum Volatility"),show.legend = TRUE)
```

``` {r fig.cap = "Weight of materials stocks by index over time\\label{fig:sector8}", fig.height=2.9, fig.width=6.5, echo=FALSE}
plot(materials1)
```

\newline
Shown below in Figure \ref{fig:sector9} are the sector weights over time for utilities stocks in both indices. The weight of utilities stocks is higher over time in the Minimum Volatility Index than in the US Equal Weight Index. This could imply that the utilities sector may not be as volatile as other sectors.

``` {r, cache=TRUE, echo=FALSE}
## Utilities
Util1 <- usa_percent[which(usa_percent$sector_name=="Utilities"), ]
Util2 <- minvol_percent[which(minvol_percent$sector_name=="Utilities"), ]
utilities1<-ggplot(Util1, aes(date, percent, colour = "US Equal Weight")) + geom_line() + 
ggtitle("Utilities Sector Weight Over Time") + xlab("Time") + 
ylab("Sector Weight") + geom_line(data = Util2, aes(x=date, y=percent, 
colour="Minimum Volatility"),show.legend = TRUE)
```

``` {r fig.cap = "Weight of utilities stocks by index over time\\label{fig:sector9}", fig.height=2.9, fig.width=6.5, echo=FALSE}
plot(utilities1)
```

\clearpage
Shown below in Figure \ref{fig:sector10} are the sector weights over time for telecommunications stocks in both indices. The weight of telecommunications stocks is higher over time in the Minimum Volatility Index than in the US Equal Weight Index. However, over time the weight of telecommunications has consistently decreased, which would imply the industry has gotten more volatile over the past few years.
``` {r, cache=TRUE, echo=FALSE}
## Telecommunication Services
Telecom1 <- usa_percent[which(usa_percent$sector_name=="Telecommunications"), ]
Telecom2 <- minvol_percent[which(minvol_percent$sector_name=="Telecommunications"), ]
telecommunication1<-ggplot(Telecom1, aes(date, percent, colour = "US Equal Weight")) + geom_line() + 
ggtitle("Telecommunication Sector Weight Over Time") + 
xlab("Time") + ylab("Sector Weight") + geom_line(data = Telecom2, aes(x=date, 
y=percent, colour="Minimum Volatility"),show.legend = TRUE)

```

``` {r fig.cap = "Weight of telecommunications stocks by index over time\\label{fig:sector10}", fig.height=3.5, fig.width=6.5, echo=FALSE}
plot(telecommunication1)
```

\clearpage 


## Trailing Volatility  
Data was collected from 10/31/2010 to 12/30/2016, from Wharton Research Data Services (WRDS) for the 908 historical constituents of the USA Equal Weight (EUSA) ETF, of which the Minimum Volatility Index is derived [@wrds2017]. Each stock's 252-day (annual) trailing volatility was calculated, and a month end spaghetti plot was produced for stocks in the Minimum Volatility Index, and the remainder of the stock comprising the US Equal Weight Index, to get a relative sense of each group's volatility attributes. The stocks comprising the Minimum Volatility Index generally had a lower volatility than those in the US Equal Weight Index. Shown below in Figure \ref{fig:vol1} are the 52-week trailing volatilities for stocks in the US Equal Weight Index ranged from 0.00492 to 87.37030, with a mean value of 1.44738.

```{r, echo = FALSE}
library(ggplot2)
load("~/thesis_final/data/monthly_data.Rda")

usa_vol <-filter(monthly_data, index_now == 0)
summary(usa_vol$volatility)

us_volplot <- ggplot(usa_vol, aes(date, volatility, group = ticker)) + geom_line() + ggtitle("52-Week Trailing Volatility for Stocks in the US Equal Weight Index") + labs(x="Date", y="52-week Trailing Volatility")
```

``` {r fig.cap = "52-week trailing volatility for stocks in the US Equal Weight Index\\label{fig:vol1}", echo=FALSE}
plot(us_volplot)
```

\clearpage 
Shown below in Figure \ref{fig:vol2} are the 52-week trailing volatilities for stocks in the Minimum Volatility Index ranged from 0.03085 to 34.20974, with a mean value of 1.40784.

```{r, echo = FALSE}
library(ggplot2)
load("~/thesis_final/data/monthly_data.Rda")

minvol_vol <-filter(monthly_data, index_now == 1)
summary(minvol_vol$volatility)

minvol_volplot <- ggplot(minvol_vol, aes(date, volatility, group = ticker)) + geom_line() + ggtitle("52-Week Trailing Volatility for Stocks in the Minimum Volatility Index") + labs(x="Date", y="52-week Trailing Volatility")
```

``` {r fig.cap = "52-week trailing volatility for stocks in the Minimum Volatility Index\\label{fig:vol2}", echo=FALSE}
plot(minvol_volplot)
```

\clearpage 

## Trailing Beta 
Data was collected from 10/31/2010 to 12/30/2016, from Wharton Research Data Services (WRDS) for the 908 historical constituents of the USA Equal Weight (EUSA) ETF, of which the Minimum Volatility Index is derived [@wrds2017]. Each stock's 252-day (annual) trailing beta was calculated, and a month end spaghetti plot was produced for stocks in the Minimum Volatility Index, and the remainder of the stock comprising the US Equal Weight Index, to get a relative sense of each group's volatility attributes. Generally, the stocks comprising the Minimum Volatility Index had a lower beta than those in the US Equal Weight Index. Shown below in Figure \ref{fig:vol3}, the 52-week trailing beta for stocks in the US Equal Weight Index ranged from -4.9037 to 6.4952, with a mean value of 1.1505.

```{r, echo = FALSE}
library(ggplot2)
load("~/thesis_final/data/monthly_data.Rda")

usa_beta <-filter(monthly_data, index_now == 0)
summary(usa_beta$beta)

usa_beta_plot <- ggplot(usa_beta, aes(date, beta, group = ticker)) + geom_line() + ggtitle("52-Week Trailing Beta for Stocks in the US Equal Weight Index") + labs(x="Date", y="52-week Trailing Beta")
```

``` {r fig.cap = "52-week trailing beta for stocks in the US Equal Weight Index\\label{fig:vol3}", echo=FALSE}
plot(usa_beta_plot)
```

\clearpage 
Shown below in Figure \ref{fig:vol4}, the 52-week trailing beta for stocks in the Minimum Volatility Index ranged from -0.2473 to 3.2021, with a mean value of 0.7856.

```{r, echo = FALSE}
library(ggplot2)
load("~/thesis_final/data/monthly_data.Rda")

minvol_beta <-filter(monthly_data, index_now == 1)
summary(minvol_beta$beta)

minvol_beta_plot <- ggplot(minvol_beta, aes(date, beta, group = ticker)) + geom_line() + ggtitle("52-Week Trailing Beta for Stocks in the Minimum Volatility Index") + labs(x="Date", y="52-week Trailing Beta")
```

``` {r fig.cap = "52-week trailing beta for stocks in the Minimum Volatility Index\\label{fig:vol4}", echo=FALSE}
plot(minvol_beta_plot)
```


